---
title: "PA1_template"
output: html_document
---

# Reproducible Research
## Peer-graded Assignment: Course Project 1
### Robert Miller
### March 12, 2017


### Reading in dataset and processing
```{r code for reading in dataset}
library(ggplot2)
library(plyr)

rawdevicedata<-read.csv("activity.csv")

devicedata<-rawdevicedata

# initial parse/validation of data
head(devicedata)
dim(devicedata)

n1<-nrow(devicedata)

# remove NA values
devicedata<-devicedata[complete.cases(devicedata),]

n2<-nrow(devicedata)

print(paste0("Rows removed from dataset: ",n1-n2,"."))

# set type for date to Date
devicedata$date<-as.Date(devicedata$date)

devicedata<-aggregate(steps~date,
                      devicedata,
                      sum)

```

### Histogram of the total number of steps taken each day
```{r histogram of total number of steps}

plot(y=devicedata$steps,x=devicedata$date,type="h",
     xlab="Date",ylab="Number of Steps",
     main="Total Number of Steps per Day Oct. 1 to Dec. 1 2012")

  
```

### Mean and Median Steps Per Day

```{r mean/median steps}

# Aggregate steps by day
sumdevicedata<-aggregate(steps ~ date, FUN=sum, data=rawdevicedata)

# output Mean and Median
meanDD<-mean(sumdevicedata$steps)
medianDD<-median(sumdevicedata$steps)

# Output Mean and Median
print(paste0("Mean number of steps across the 61 day period: ", meanDD, "."))
print(paste0("Median number of steps across the 61 day period: ", medianDD, "."))

```

### Average Steps Time Series Plot
### Time series plot of the average number of steps taken
```{r time series plot}
meandevicedata<-aggregate(steps~date,FUN=mean,data=rawdevicedata)

meandevicedata$date<-as.Date(meandevicedata$date)

plot(meandevicedata,type="l",ylab="Mean Steps",
     xlab="Date",
     main="Mean Steps Per Day")
```

### The 5-minute interval that, on average contains the maximum number of steps
```{r}
maxdates<-aggregate(steps~date,rawdevicedata,max)

addInterval<-merge(maxdates,rawdevicedata)

avgInterval<-as.integer(mean(addInterval$interval))
print(paste0("Mean 5-minute interval that contains max steps: ", avgInterval, "."))

```
### Data Imputation
```{r data imputation}

# Assign raw dataset to temp dataset
tempdevicedata<-rawdevicedata

# Divide mean steps per day by number of intervals
# assign to overwrite NA values
tempdevicedata<-tempdevicedata[is.na(rawdevicedata$steps),]

tempdevicedata$steps<-meanDD/288


# carve out non-NA cases
completecases<-rawdevicedata[complete.cases(rawdevicedata), ]



# recombine imputed rows and complete case rows
tempdevicedata<-rbind(completecases,tempdevicedata)

# recompute mean, including imputed days
meandevicedata<-aggregate(steps~date,FUN=mean,data=tempdevicedata)

# converte date feature to enable clean plot
meandevicedata$date<-as.Date(meandevicedata$date)


plot(meandevicedata,type="l",ylab="Mean Steps",
     xlab="Date",
     main="Mean Steps Per Day with Imputed Values")
```

### Histogram with Imputed Data
```{r new histogram with imputed data}

histdata<-aggregate(steps~date,
                      tempdevicedata,
                      sum)

histdata$date<-as.Date(histdata$date)

plot(y=histdata$steps,x=histdata$date,type="h",
     xlab="Date",ylab="Number of Steps",
     main="Total Number of Steps per Day Oct. 1 to Dec. 1 2012 - Imputed")

```

### Comparison of Mean Steps Per Day by Weekday and Weekend
```{r}

wkndList<-c('Saturday','Sunday')


tempdevicedata$date<-as.Date(tempdevicedata$date)

tempdevicedata$wk<-factor((weekdays(tempdevicedata$date) %in% wkndList), levels = c(TRUE, FALSE),
                    labels=c('weekend','weekday'))


wknddevicedata<-tempdevicedata[tempdevicedata$wk=="weekend",]

wkdydevicedata<-tempdevicedata[tempdevicedata$wk!="weekend",]


wknddevicedata<-aggregate(steps~date,
                          wknddevicedata,
                          mean)


par(mfrow=c(1,2))
plot(wknddevicedata,type="l",ylab="Mean Steps",
     xlab="Date",
     ylim=c(0,80),
     main="Weekend - Mean Steps Per Day")

wkdydevicedata<-aggregate(steps~date,
                          wkdydevicedata,
                          mean)


plot(wkdydevicedata,type="l",ylab="Mean Steps",
     xlab="Date",
     ylim=c(0,80),
     main="Weekday - Mean Steps Per Day")



```