---
title: "RepData_PeerAssessment1"
output: html_document
---

# Reproducible Research
## Peer-graded Assignment: Course Project 1
### Robert Miller
### March 7, 2017


1. Code for reading int he dataset and/or processing the data
```{r code for reading in dataset}
library(ggplot2)
library(plyr)

rawdevicedata<-read.csv("activity.csv")

devicedata<-rawdevicedata

# initial parse/validation of data
head(devicedata)
dim(devicedata)

n1<-nrow(devicedata)

# remove NA values
devicedata<-devicedata[complete.cases(devicedata),]

n2<-nrow(devicedata)

# post processing parse/validation of data
head(devicedata)
dim(devicedata)

print(paste0("Rows removed from dataset: ",n1-n2,"."))

# set type for date to Date
devicedata$date<-as.Date(devicedata$date)

```

2. Histogram of the total number of steps taken each day
```{r histogram of total number of steps}

plot(y=devicedata$steps,x=devicedata$date,type="h",
     xlab="Date",ylab="Number of Steps",
     main="Total Number of Steps per Day Oct. 1 to Dec. 1 2012")

  
```

3. Mean and Median Steps Per Day

```{r mean/median steps}

# Aggregate steps by day
sumdevicedata<-aggregate(steps ~ date, FUN=sum, data=rawdevicedata)

# output Mean and Median
meanDD<-mean(sumdevicedata$steps)
medianDD<-median(sumdevicedata$steps)

# Output Mean and Median
print(paste0("Mean number of steps across the 61 day period: ", meanDD, "."))
print(paste0("Median number of steps across the 61 day period: ", medianDD, "."))

```

### Average Steps Time Series Plot
4. Time series plot of the average number of steps taken
```{r time series plot}
meandevicedata<-aggregate(steps~date,FUN=mean,data=rawdevicedata)

meandevicedata$date<-as.Date(meandevicedata$date)

plot(meandevicedata,type="l",ylab="Mean Steps",
     xlab="Date",
     main="Average Number of Steps Per Day")
```

5. The 5-minute interval that, on average contains the maximum number of steps
```{r}
maxdates<-aggregate(steps~date,rawdevicedata,max)

addInterval<-merge(maxdates,rawdevicedata)

avgInterval<-as.integer(mean(addInterval$interval))
print(paste0("Mean 5-minute interval that contains max steps: ", avgInterval, "."))

```
### Data Imputation
```{r data imputation}

tempdevicedata<-rawdevicedata
tempdevicedata[is.na(rawdevicedata$steps),"steps"]<-meanDD/288

rawdevicedata<-complete.cases(rawdevicedata)

tempdevicedata<-rbind(tempdevicedata,rawdevicedata)

meandevicedata<-aggregate(steps~date,FUN=mean,data=tempdevicedata)

meandevicedata$date<-as.Date(meandevicedata$date)


plot(meandevicedata,type="l",ylab="Mean Steps",
     xlab="Date",
     main="Average Number of Steps Per Day")
```
```

### Nacho Cheese Doritos